---
title: "lab04"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lab04}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include = FALSE}
library(lab04)
```


## Introduction

The `lab04` package has one main function, `linreg()`. It fits multiple
linear regression models, which estimates using the QR decomposition,
and returns an S3 object of class `linreg`. The object works with the
usual methods from base R, such as `print()`, `summary()`, `coef()`,
`resid()`, `pred()` and `plot()`.

The examples in this vignette use the `iris` dataset, which is included
in R. The dataset contains measurements of 150 flowers from three
species (*setosa*, *versicolor* and *virginica*).

The vignette first describes how `linreg()` works. It then fits a model and shows each of the methods that can be used on the resulting object.

## The linreg() function

### Usage

`linreg(formula, data)`

- `formula`: a formula of the form `y ~ x1 + x2 + ...`, the same as for
  `lm()`.
- `data`: a data frame containing the variables in the formula.

The function first checks that `formula` is a formula and that `data` is
a data frame, and stops with an error message if they are not.

### How the model is estimated

The design matrix $\mathbf{X}$ is created from the formula with
`model.matrix()`, and the response $\mathbf{y}$ is taken from the data
using `all.vars()`. The model is

$$\mathbf{y} = \mathbf{X}\beta + \varepsilon.$$

Instead of calculating $(\mathbf{X}^T\mathbf{X})^{-1}$ directly,
`linreg()` use QR decomposition $\mathbf{X} = \mathbf{Q}\mathbf{R}$,
where $\mathbf{Q}$ is orthogonal
($\mathbf{Q}^T\mathbf{Q} = \mathbf{I}$), and $\mathbf{R}$ is upper
triangular. This is seen as numerically more stable and is also what
`lm()` does.

**Coefficients;** Since
$\mathbf{X}^T\mathbf{X} = \mathbf{R}^T\mathbf{R}$, the normal equations
reduce to

$$\mathbf{R}\hat\beta = \mathbf{Q}^T\mathbf{y}$$

This can be solved with back-substitution as matrix $\mathbf{R}$ is a
triangular matrix.

**Fitted values and residuals;**

$$\hat{\mathbf{y}} = \mathbf{X}\hat\beta, \qquad \hat{\mathbf{e}} = \mathbf{y} - \hat{\mathbf{y}}$$

**Degrees of freedom and residual variance;** With $n$ observations and
$p$ parameters,

$$df = n - p, \qquad \hat\sigma^2 = \frac{\hat{\mathbf{e}}^T\hat{\mathbf{e}}}{df}.$$

**Variance of the coefficients;** Using QR decomposition,
$(\mathbf{X}^T\mathbf{X})^{-1}$ becomes
$\mathbf{R}^{-1}(\mathbf{R}^{-1})^T$, thus,

$$\widehat{\text{Var}}(\hat\beta) = \hat\sigma^2 \, \mathbf{R}^{-1}(\mathbf{R}^{-1})^T,$$

where the diagonal of this matrix gives the variance of the
coefficients.

**t-values and p-values;**

$$t_{\beta} = \frac{\hat\beta}{\sqrt{\widehat{\text{Var}}(\hat\beta)}}$$

The two-sided p-values are calculated from the $t$-distribution with
$df$ degrees of freedom using `pt()`.

## Fitting a model

```{r}
mod <- linreg(Petal.Length ~ Species, data = iris)
class(mod)
names(mod)
```

## Methods

The package includes the methods `print`, `summary`, `coef`, `resid`,
`pred`, and `plot.` The metod functions import necessary fucntions from
the `stats` and `ggplot2` packages.

### print()

Prints the regression coefficients of a linreg object.

#### Usage

`print(x)`

- `x` : An object of class linreg
- `...` : Optional additional arguments for compatibility with generic
  method

`print` example:

```{r}
print(mod)
```

### summary()

Prints a summary of the regression results for a linreg object.
Invisibly returns the linreg object.

#### Usage

`summary(object)`

- `object` : An object of class linreg
- `...` : Optional additional arguments for compatibility with generic
  method

`summary` example:

```{r}
summary(mod)
```

### coef()

Returns a named vector of regression coefficients of a linreg object.

#### Usage

`coef(object)`

- `object` : An object of class linreg
- `...` : Optional additional arguments for compatibility with generic
  method

`coef` example:

```{r}
coef(mod)
```

### resid(), residuals()

Returns an array of linear regression residuals as the difference
between predicted and actual response values. `resid()` and
`residuals()` are synonyms.

#### Usage

`resid(object)`  
`residuals(object)`

- `object` : An object of class linreg
- `...` : Optional additional arguments for compatibility with generic
  method

`resid` example (output here limited to the 5 first lines):

```{r}
resid(mod) |> 
  head(5)
```

### pred()

Returns an array of the predicted values for a linreg object.

#### Usage

`pred(object)`

- `object` : An object of class linreg
- `...` : Optional additional arguments for compatibility with generic
  method

`pred` example (output here limited to the 5 first lines):

```{r}
pred(mod) |> 
  head(5)
```

### plot()

Prints two diagnostic plots for a linreg object. Invisibly returns a
list with two ggplot2 objects.

#### Usage

`plot(x)`

- `x` : An object of class linreg
- `...` : Optional additional arguments for compatibility with generic
  method

`plot` example:

```{r}
plot(mod)
```
