---
title: "lab04"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lab04}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Installing and loading the package

The package can be installed using

```{r, results = 'hide', message=FALSE}
pak::pkg_install("https://github.com/karinalfrida/advanced-r-lab04")
```

and then, when successfully installed, be loaded

```{r}
library(lab04)
```

```{r, include = FALSE}
mod <- linreg(Petal.Length ~ Species, data = iris)
```


## Methods

The package includes the methods `print`, `summary`, `coef`, `resid`, `pred`, and `plot.`
The metod functions import necessary fucntions from the `stats` and `ggplot2` packages.

### print()

Prints the regression coefficients of a linreg object.

#### Usage

```{r, eval = FALSE}
print(x)
```

- `x` : An object of class linreg
- `...` : Optional additional arguments for compatibility with generic method

`print` example:

```{r}
print(mod)
```

### summary()

Prints a summary of the regression results for a linreg object.
Invisibly returns the linreg object.

#### Usage

```{r, eval = FALSE}
summary(object)
```

- `object` : An object of class linreg
- `...` : Optional additional arguments for compatibility with generic method

`summary` example:

```{r}
summary(mod)
```

### coef()

Returns a named vector of regression coefficients of a linreg object.

#### Usage

```{r, eval = FALSE}
coef(object)
```

- `object` : An object of class linreg
- `...` : Optional additional arguments for compatibility with generic method

`coef` example:

```{r}
coef(mod)
```

### resid(), residuals()

Returns an array of linear regression residuals as the difference between predicted and actual response values. `resid()` and `residuals()` are synonyms.


#### Usage

```{r, eval = FALSE}
resid(object)
residuals(object)
```


- `object` : An object of class linreg
- `...` : Optional additional arguments for compatibility with generic method

`resid` example (output here limited to the 10 first lines):

```{r}
resid(mod) |> 
  head(10)
```

### pred()

Returns an array of the predicted values for a linreg object.

#### Usage

```{r, eval = FALSE}
pred(object)
```

- `object` : An object of class linreg
- `...` : Optional additional arguments for compatibility with generic method

`pred` example (output here limited to the 10 first lines):

```{r}
pred(mod) |> 
  head(10)
```

### plot()

Prints two diagnostic plots for a linreg object.
Invisibly returns a list with two ggplot2 objects.

#### Usage

```{r, eval = FALSE}
plot(x)
```

- `x` : An object of class linreg
- `...` : Optional additional arguments for compatibility with generic method

`plot` example:

```{r}
plot(mod)
```
