---
title: "lab04"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lab04}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lab04)
```

## Introduction

The `lab04` package has one main function, `linreg()`. It fits multiple linear
regression models, which estimates using the QR decomposition, and returns an
S3 object of class `my_linreg`. The object works with the usual methods
from base R, such as `print()`, `summary()`, `coef()`, `resid()`, `pred()` and `plot()`.

The examples in this vignette use the `iris` dataset, which is included in R.
The dataset contains measurements of 150 flowers from three species (*setosa*,
*versicolor* and *virginica*).

The vignette first describes how `linreg()` works. It then fits a model and
shows each of the methods that can be used on the result.

## The linreg() function

### Usage

```r
linreg(formula, data)
```

- `formula`: a formula of the form `y ~ x1 + x2 + ...`, the same as for `lm()`.
- `data`: a data frame containing the variables in the formula.

The function first checks that `formula` is a formula and that `data` is a data
frame, and stops with an error message if they are not.

### How the model is estimated

The design matrix $\mathbf{X}$ is created from the formula with
`model.matrix()`, and the response $\mathbf{y}$ is taken from the data using
`all.vars()`. The model is

$$\mathbf{y} = \mathbf{X}\beta + \varepsilon.$$

Instead of calculating $(\mathbf{X}^T\mathbf{X})^{-1}$ directly, `linreg()` use
QR decomposition $\mathbf{X} = \mathbf{Q}\mathbf{R}$, where $\mathbf{Q}$ is
orthogonal ($\mathbf{Q}^T\mathbf{Q} = \mathbf{I}$), and $\mathbf{R}$ is upper
triangular. This is seen as numerically more stable and is also what `lm()` does.

**Coefficients;**
Since $\mathbf{X}^T\mathbf{X} = \mathbf{R}^T\mathbf{R}$, the
normal equations reduce to

$$\mathbf{R}\hat\beta = \mathbf{Q}^T\mathbf{y}$$

This can be solved with back-substitution as matrix $\mathbf{R}$ is a
triangular matrix.

**Fitted values and residuals;**

$$\hat{\mathbf{y}} = \mathbf{X}\hat\beta, \qquad \hat{\mathbf{e}} = \mathbf{y} - \hat{\mathbf{y}}$$

**Degrees of freedom and residual variance;** With $n$ observations and $p$
parameters,

$$df = n - p, \qquad \hat\sigma^2 = \frac{\hat{\mathbf{e}}^T\hat{\mathbf{e}}}{df}.$$

**Variance of the coefficients;** Using QR decomposition,
$(\mathbf{X}^T\mathbf{X})^{-1}$ becomes $\mathbf{R}^{-1}(\mathbf{R}^{-1})^T$, thus,

$$\widehat{\text{Var}}(\hat\beta) = \hat\sigma^2 \, \mathbf{R}^{-1}(\mathbf{R}^{-1})^T,$$

where the diagonal of this matrix gives the variance of the coefficients.

**t-values and p-values;**

$$t_{\beta} = \frac{\hat\beta}{\sqrt{\widehat{\text{Var}}(\hat\beta)}}$$

The two-sided p-values are calculated from the $t$-distribution with $df$
degrees of freedom using `pt()`.

## Fitting a model


```{r}
mod <- linreg(Petal.Length ~ Species, data = iris)
class(mod)
names(mod)
```

## Methods

### print()

### summary()

### coef()

### resid()

### pred()

### plot()


