---
title: "lab04"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lab04}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lab04)
```

## Introduction

fits multiple linear regression models. It

The `lab04` package has one main function, `linreg()`. It fits multiple linear
regression models, which estimates using the QR decomposition, and returns an
S3 object of class `linreg` !!(input class name wr agree on)!!. The object works with the usual methods
from base R, such as `print()`, `summary()`, `coef()`, `resid()`, `pred()` and `plot()`.

The examples in this vignette use the `iris` dataset, which is included in R.
The dataset contains measurements of 150 flowers from three species (*setosa*,
*versicolor* and *virginica*).

The vignette first describes how `linreg()` works. It then fits a model and
shows each of the methods that can be used on the result.

## The linreg() function

### Usage

```r
linreg(formula, data)
```



## Fitting a model

TODO

```{r}
mod <- linreg(Petal.Length ~ Species, data = iris)
class(mod)
names(mod)
```

## Methods

### print()

### summary()

### coef()

### resid()

### pred()

### plot()


